---
title: "Stochastic assembly"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stochastic_assembly}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette provides an example of running a stocahstic assembly use plant.assembly. Below we consider both a 1 and 2 trait example.

First, basic setup:

```{r, setup}
library(plant)
library(plant.assembly)

library(tidyverse)
library(ggplot2)

# devtools::load_all()
#devtools::load_all("../plant-dev")

# logging of output
plant_log_console()

# set controls on plant package 
plant_control <- function() {
  ctrl <- scm_base_control()
  ctrl$equilibrium_nsteps <- 80
  ctrl$equilibrium_solver_name <- "iteration"
  ctrl
}

# set baseline parameters
p <- scm_base_parameters("FF16")
p$strategy_default$a_l1 <- 2.17
p$strategy_default$a_l2 <- 0.5
p$max_patch_lifetime <- 60

```

# One trait: lma

Now let's run the assembly for a single trait

## Assembly

```{r}
# Set absolute bounds
bounds0 <- bounds(lma = c(0.01, 1))

# set controls on assembly
control <- assembler_control(
  list(
    run_type = "single",
    birth_type = "stochastic",
    vcv = diag(1)
  )
)

# setup the assembler to start with empty community
community0 <- community_start(p, bounds0)
obj <- assembler_start(community0, control = control)

# run assembler
obj <- assembler_run(obj, nsteps = 100)

saveRDS(obj, "output/stochastic.rds")
```

## Analysis

```{r}
obj <- readRDS("output/stochastic-lma-2023.10.24-fitness.rds")
```

Tidy an assembly object. This extracts and tidies the community (i.e endpoint-community) resident traits, the community fitness landscape (i.e endpoint-fitness landscape), and the historical timeseries of the two properties above. 
```{r}
tidied_assembly <- tidy_assembly(obj)
```

Communities (a slice through time) use the same plotting function as historical community timeseries

```{r}
community_plot <- plot_community(tidied_assembly$community)
history_plots <- plot_community(tidied_assembly$history)
```

## Fitness landscapes

Plot fitness landscapes. These also use the same function for both communities and historical timeseries of communities.

```{r}
community_landscape_plots <- plot_landscape(tidied_assembly$community_landscape)
history_landscape_plots <- plot_landscape(tidied_assembly$history_landscape)

imap(history_landscape_plots, ~ggsave(paste0("output/stochastic/stochastic_plot_history_landscape_one_trait/",.y,".png"), .x))
```


# Two traits: lma + hmat

Now let's run the assembly for two traits

## Assembly

```{r}

bounds0 <- bounds(lma = c(0.01, 1), hmat = c(0.5, 30))

# controls on assembly
control <- assembler_control(
  list(
    run_type = "single",
    birth_type = "stochastic",
    compute_viable_fitness = FALSE,
    vcv = diag(2)
  )
)

# setup the assembler to start with empty community
community0 <- community_start(p, bounds0)
obj <- assembler_start(community0, control = control)

# run assembly
obj <- assembler_run(obj, nsteps =10)
```

```{r}
saveRDS(obj, "output/stochastic_two_species.rds")
```

```{r}
obj <- readRDS("output/stochastic_two_species.rds")
```

```{r}
tidied_assembler <- tidy_assembly(obj)
```

## Analysis

Now lets' look at the results


Plot communities (historical and latest) for two traits

```{r}
community_plot <- plot_community_2d(tidied_assembler$community)
history_plots <- plot_community_2d(tidied_assembler$history)
```


How many species in the community?
```{r}
p$max_patch_lifetime
p$strategy_default$hmat
obj$community$traits
obj$community$birth_rate
```

```{r}
landscape <- obj$community$fitness_approximate_points %>% as_tibble()
ggplot(landscape, aes(lma, fitness)) + 
  geom_line() +
  scale_x_log10() +
  geom_point(data = tibble(lma = obj$community$traits[,1], fitness = 0), col="red")
```

