---
title: "max_fitness assmebly"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{max_fitness}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


This vignette provides an example of running community assembly using the max_fitness method in  plant.assembly. Below we consider both a 1 and 2 trait example.

First, basic setup:

# Example of assembly for lma, using maxfit 


```{r, setup}
library(plant)
library(plant.assembly)
devtools::load_all()
# devtools::load_all("../plant-dev")

# logging of output
plant_log_console()

# set baseline parameters
p <- plant_default_assembly_pars()

model_support <- list(
  p = plant_default_assembly_pars(),
  plant_control = plant_default_assembly_control()
)

```


# One trait: lma

Now let's run the assembly for a single trait

## Assembly

```{r}
# Set absolute bounds
bounds0 <- bounds(lma = c(0.01, 2))

# set controls on assembly
control <- assembler_control(
  list(
    run_type = "to_equilibrium",
    birth_type = "maximum",
    birth_move_tol = 1,
    compute_viable_fitness = FALSE
  )
)

# setup the assembler to start with empty community
community0 <- 
  community_start(bounds0,
    fitness_control = list(type = "grid", n = 10),
    model_support = model_support)

obj <- assembler_start(community0, control = control)

# run assembler
obj <- assembler_run(obj, nsteps = 20)
```

Save output
```{r}
saveRDS(obj, "output/fitmax-lma.rds")

```

## Analysis
Now lets' look at the results

```{r}
obj <- readRDS("output/fitmax-lma-old1.rds")
```

How many species in the community?
```{r}
obj$community$traits
obj$community$birth_rate
```

```{r}
library(dplyr)
library(ggplot2)
landscape <- obj$community$fitness_points

ggplot(landscape, aes(lma, fitness)) + 
  geom_line() +
  scale_x_log10() +
  geom_point(data = tibble(lma = obj$community$traits[,1], fitness = approx(landscape$lma, landscape$fitness,  obj$community$traits[,1])$y), col="red")
```

Generate a history of fitness landscapes across steps, as species are introduced and removed.

```{r}
assembly <- tidy_assembly(obj)

fitmax_plots <- plot_landscape(assembly$history_landscape)
```


# One trait: Height

## Assembly

```{r}
# Set absolute bounds
bounds0 <- bounds(hmat = c(0.5, 30))

# set controls on assembly
control <- assembler_control(
  list(
    run_type = "to_equilibrium",
    birth_type = "maximum",
    birth_move_tol = 1,
    compute_viable_fitness = FALSE
  )
)

# setup the assembler to start with empty community
community0 <- 
  community_start(bounds0,
    fitness_control = list(type = "gp"),
     model_support = model_support)

obj <- assembler_start(community0, control = control)

# run assembler
obj <- assembler_run(obj, nsteps = 20)

saveRDS(obj, "output/fitmax-hmat.rds")
```

# Appendix

Common mixtures of lma for this distrubance regime

         lma birth_rate
1 0.03886086   277.2675
2 0.19630583   18.42347
3 0.11070529   67.25629
4 0.25085100   2.852756
