---
title: "Solving for attarctors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{max_fitness}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette provides an example of calculating attractors for 1 spp communities, using 1,2 or more traits. 


First, some basic setup

```{r, setup}
library(plant)
library(plant.assembly)
library(parallel)
library(tidyverse)

devtools::load_all()
# devtools::load_all("../plant-dev")

# logging of output
plant_log_console()

# set controls on plant package 
plant_control <- function() {
  ctrl <- scm_base_control()
  ctrl$equilibrium_nsteps <- 20
  ctrl$equilibrium_solver_name <- "hybrid"
  ctrl
}

# set baseline parameters
p <- scm_base_parameters("FF16")
p$strategy_default$a_l1 <- 2.17
p$strategy_default$a_l2 <- 0.5
p$strategy_default$hmat <- 10
p$max_patch_lifetime <- 60


```

# Example for 1 trait (lma)

## Selection gradients

Solving for the 1D attarctor is possible with the community objects. 

Let's say you want to calculate the selection gradient at the following trait value. 

```{r}
trait <- 0.04045

# setup the assembler to start with empty community
community0 <-
  community_start(
    p,
    bounds(lma = c(0.01, 2)),
    fitness_approximate_control = list(type = "grid")
  )

community_eq <-
  # Empty community
  community0 %>%
  # Add a startegy specified by it's trait value
  community_add(plant::trait_matrix(trait, "lma")) %>%
  # solve euqilibrium density
  community_run_to_equilibrium() %>%
  # calculate fitness gradient
  community_selection_gradient()

community_eq$fitness
community_eq$selection_gradient
```

In this example fitness is near zero (as it should be), because the community is at euqilibrium. The selection gradient is positive meanign, selection is to increase trait value. 

Let's visualise this by calculating a fitness landscape:

```{r}
community_eq <- community_eq %>%
  community_prepare_approximate_fitness()

data <- 
  community_eq$fitness_approximate_points %>% as_tibble() %>% 
  mutate(resident = ifelse(lma %in% community_eq$traits, TRUE, FALSE))

ggplot(data, aes(lma, fitness)) +
  geom_line() +
  geom_point(data = filter(data, resident), col="red") +
  scale_x_log10()
```

So let's plot the selection gradient across a range of trait values. It's clear from the above that the attractor lies in the range 0.04-0.08.

I'll make a worker function that runs the above routines. 

```{r}
f <- function(x, community) {
  community %>%
    community_add(plant::trait_matrix(x, "lma")) %>%
    community_run_to_equilibrium() %>%
    community_selection_gradient() 
}

# Create a data frame with range of trait values and solve

data_communities <- 
  # trait matix
  tibble(traits = seq_log_range(c(0.03, 0.05), length.out = 5)) %>%
  mutate(
    # solve commuities
    communities = parallel::mclapply(
      traits, function(x) f(x, community0),
      mc.cores = detectCores() - 2
    ),
    # extract gradient
    selection_gradient = map_dbl(communities, ~ .x$selection_gradient)
  )

# plot
plot_sg <- 
  ggplot(data_communities, aes(traits, selection_gradient)) +
  geom_line(col="red") +
  geom_point(col="red", pch="x") +
  geom_abline(intercept = 0, linetype = "dashed") +
  ylim(-10000, 10000) +
  scale_x_log10()

plot_sg
```

## Using a root solver to find the attractor

We've wrapped up the above routines to enable solving of the 1D attractor

```{r}
community_attractor <- community0 %>%
 community_solve_singularity_1D(bounds = c(0.035, 0.045), tol = 1e-4)

community_attractor$traits
```

Let's add this to our previous plot
```{r}
plot_sg +
  geom_point(col="red",
    data = tibble(
      traits = community_attractor$traits, selection_gradient = community_attractor$selection_gradient 
    )
  )
```


# 2D Example: Attarctor for lma and hmat

## Selection gradients

Solving for the 2D attarctor is possible with the community objects. 

Let's say you want to calculate the selection gradient at the following trait values. 

```{r}
traits <- c(0.04045, 10)

# setup the assembler to start with empty community
community0 <-
  community_start(
    p,
    bounds(lma = c(0.01, 2), hmat = c(0.01, 100)),
    fitness_approximate_control = list(type = "grid")
  )

community_eq <-
  # Empty community
  community0 %>%
  # Add a startegy specified by it's trait value
  community_add(plant::trait_matrix(traits, c("lma", "hmat"))) %>%
  # solve euqilibrium density
  community_run_to_equilibrium() %>%
  # calculate fitness gradient
  community_selection_gradient()

community_eq <- community_eq %>% community_selection_gradient()

community_eq$fitness
community_eq$selection_gradient
```

In this example fitness is near zero (as it should be), because the community is at euqilibrium. The selection gradient is positive meanign, selection is to increase trait value. 

So let's plot the selection gradient across a range of trait values. It's clear from the above that the attractor lies in the range 0.04-0.08.

I'll make a worker function that runs the above routines. 

```{r}
f <- function(x, community) {
  community %>%
    community_add(plant::trait_matrix(x, c("lma", "hmat"))) %>%
    community_run_to_equilibrium() %>%
    community_selection_gradient() 
}

# Create a data frame with range of trait values and solve

data_communities <- 
  # trait matix
  expand_grid(
    lma = seq_log_range(c(0.03, 0.05), length.out = 5),
    hmat = seq_log_range(c(5, 15), length.out = 5)
  ) %>%
  mutate(i = seq_len(n())) %>%
  nest(.by = i, .key = "traits") %>%
  mutate(
    # solve commuities
    communities = parallel::mclapply(
      traits, function(x) f(as.matrix(x), community0),
      mc.cores = detectCores() - 2
    ))

data_communities2 <-
  data_communities %>%
    mutate(
      selection_gradient_lma = map_dbl(communities, ~ .x$selection_gradient[1]),
      selection_gradient_hmat = map_dbl(communities, ~ .x$selection_gradient[2])
    ) %>% 
    unnest(traits)
    
# plot - DIY

plot_sg <- 
  data_communities2 %>% 
    select(lma, hmat, selection_gradient_lma, selection_gradient_hmat) %>%
    ggplot(aes(x = lma, y = hmat, 
      xend = lma + rescale(selection_gradient_lma, 0.02/4), 
      yend = hmat + rescale(selection_gradient_hmat, 10/10))) +
  geom_point() +
  geom_segment(linewidth = 0.25, arrow = arrow(length = unit(0.01, "npc")))

# ggquiver is deisgned for this 
# but only works when x axes are log transformed
# could add custom axis labe;l back to original values?
library(ggquiver)

 rescale <- function(x, s = 1) {x / max(abs(range(x))) * s }

plot_sg <-
  data_communities2 %>% 
    ggplot(aes(x = log(lma), y = log(hmat), u = selection_gradient_lma, v= selection_gradient_hmat)) +
    geom_point() +
    geom_quiver(vecsize=NULL)

plot_sg
```

## Using a root solver to find the attractor

We've wrapped up the above routines to enable solving of the 1D attractor

```{r}
community_attractor <- community0 %>%
 community_solve_singularity_1D(bounds = c(0.035, 0.045), tol = 1e-4)

community_attractor$traits
```

Let's add this to our previous plot
```{r}
plot_sg +
  geom_point(col="red",
    data = tibble(
      traits = community_attractor$traits, selection_gradient = community_attractor$selection_gradient 
    )
  )
```

